---
title: "hw3_yg2625"
author: "Yue Gu"
date: "October 13, 2018"
output: github_document
---


```{r setup, include=FALSE}
library(tidyverse)
library(p8105.datasets)

knitr::opts_chunk$set(echo = TRUE,
                      fig.width = 15,
                      fig.height = 5,
                      fig.asp = 0.6,
                      out.width = "90%")

```



# Problem 1
## Data Cleaning
```{r}
data(brfss_smart2010)
brfss_data = 
  brfss_smart2010 %>% 
  janitor::clean_names() %>% 
  rename(.data = ., state = locationabbr, county = locationdesc) %>% 
  filter(.data = ., topic == "Overall Health", response == "Excellent" || response == "Very good" || response == "Good" || response == "Fair" || reponse == "Poor") %>% 
  mutate(.data = ., response = as.factor(response)) %>% 
  arrange(.data = .,  match(response,c("Excellent", "Very good", "Good", "Fair", "Poor")))

```

## Answering questions
```{r}
brfss_data %>% 
  filter(year == 2002) %>% 
  group_by(state) %>% 
  summarize(n_obs = n_distinct(county) == 7) %>% 
  filter(n_obs == T)
```
Hence, by the output, CT, FL, NC were observed at 7 locations in 2002.

The spaghetti plot shown below
```{r}
brfss_data %>% 
  group_by(year, state) %>% 
  summarize(n_obs = n_distinct(county)) %>%
  ggplot(aes(x = year, y = n_obs, color = state)) + geom_line()
```

Required table shown below
```{r}
brfss_data %>% 
  filter(.data = ., response == "Excellent", state == "NY", year == "2002"|year == "2006"|year == "2010") %>%
  group_by(year, state) %>% 
  summarize(mean_data = mean(data_value)，
            sd_data = sd(data_value)) %>% 
  knitr::kable(digits = 2)
```

Required five-panel plot shown below
```{r}
brfss_data %>% 
  group_by(year, state, response) %>% 
  summarize(mean_data = mean(data_value)) %>% 
  ggplot(aes(x = year, y = mean_data, color = state)) + 
    geom_line() + 
    facet_grid(.~response)
```

# Problem 2

## Description of exploration
```{r}
data("instacart")
insta_data =
  instacart
  

dim(insta_data)
n_distinct(insta_data$order_id)
n_distinct(insta_data$product_id)
n_distinct(insta_data$user_id)
n_distinct(insta_data$aisle_id)
n_distinct(insta_data$department_id)
```

## Aisles
```{r}
n_distinct(insta_data$aisle)
head(arrange(count(insta_data, aisle_id), desc(n)), n = 1)

## Plot
insta_data %>% 
  group_by(aisle_id) %>% 
  summarize(n = n()) %>% 
  ggplot(aes(x = as.character(aisle_id), y = n, color = as.character(aisle_id))) +
    geom_point(alpha = 0.5) +
    viridis::scale_color_viridis(
    name = "Aisle ID", 
    discrete = TRUE
  ) + 
    theme(legend.position = "bottom")
```

## Table for most pop items
```{r}
insta_data %>% 
  filter(.data = ., aisle == "baking ingredients" | aisle == "dog food care" | aisle == "packaged vegetables fruits") %>% 
  group_by(aisle, product_name) %>%
  summarize(n = n()) %>% 
  filter(min_rank(desc(n)) == 1) %>% 
  knitr::kable()
```

## Table for hr of the day
```{r}
insta_data %>% 
  filter(.data = ., product_name == "Pink Lady Apples" | product_name == "Coffee Ice Cream") %>% 
  group_by(order_dow, product_name) %>% 
  summarize(mean_hr = mean(order_hour_of_day)) %>% 
  spread(key = order_dow, value = mean_hr) %>%
  rename(Sun = "0", Mon = "1", Tues = "2", Wed = "3", Thrs = "4", Fri = "5", Sat = "6", Product = product_name) %>% 
  knitr::kable(digit = 2)
```

# Problem 3
## Data exploration

```{r}
data(ny_noaa)
noaa_data = 
  ny_noaa

dim(ny_noaa)
sum(is.na(ny_noaa$prcp))
sum(is.na(ny_noaa$snow))
sum(is.na(ny_noaa$snwd))
sum(is.na(ny_noaa$tmax))
sum(is.na(ny_noaa$tmin))

```

## Data cleaning
```{r}
noaa_data_tidy =
  noaa_data %>% 
  separate(date, into = c("year", "month", "day"), sep = "-") %>% 
  mutate(.data = ., prcp = as.numeric(prcp)/10, tmax = as.numeric(tmax)/10, tmin = as.numeric(tmin)/10)


head(arrange(count(noaa_data_tidy, snow), desc(n)), n = 1)
```
0 is the most commonly obeserved value because not every day is snowing and when there's 0 snowfall when no snow is falling and the proportion of days in a year is relatively high.

## Two-panel plot for avg max temp
```{r}
noaa_data_tidy %>% 
  filter(.data = ., month == "01" | month == "07", tmax != "") %>% 
  group_by(id, month, year) %>%
  summarize(tmax_mean = mean(tmax)) %>% 
  ggplot(aes(x = year, y = tmax_mean)) +
    geom_point()+
    facet_grid(. ~ month)
```


## Two-panel plot for last question
```{r}
library(patchwork)
t_vs_p = ggplot(noaa_data_tidy, aes(x = tmax, y = tmin)) +
  geom_hex()

noaa_data_snow =
  noaa_data_tidy %>% 
  filter(.data = ., snow > 0 & snow < 100)

snow_dist_p = ggplot(noaa_data_snow, aes(x = snow, fill = year)) +
  geom_density()

t_vs_p + snow_dist_p

```









